Think deep for reef resilience

### 2025-08-08 paper/paper_draft.qmd
Expand depth estimation method

Median depths for each reef location were extracted from satellite derived bathymetry data
for the Great Barrier Reef (GBRMPA [-@gbr10gb2020]) (see discussion for more details). Due
to varying degrees of alignment between bathymetry data and the defined outlines for GBR
reefs, reefs with less than 5% coverage of valid bathymetry data were removed from analyses
involving depth as a resilience factor.

\[**TI:** **For Takuya to flesh out**\]

\[Below moved from discussion - to be reworked for flow\]

These data (GBRMPA [-@gbr10gb2020]) are derived from satellite imagery and can be limited in
their alignment with the established GBRMPA reef outlines used in the investigation to
characterise reef locations. As a result, 842 reefs (22% of 3806 reefs) were removed from
the depth-related analyses due to having less than 5% coverage of depth data. **\[TI to BG:
is this number of reefs consistent?\]**

Additionally, deriving estimates of depth from satellite imagery is prone to high degrees of
uncertainty. Data adopted for the given assessment are said to be accurate up to 15m from
Mean Sea Level [@gbr10gb2020]. Furthermore, an estimated median depth is determined for each
reef, using the established GBRMPA reef outlines to provide an indication of “true” reef
extent. Therefore, the estimates of reef depths are taken as a useful indication rather than
an accurate and precise report of reef depths across the GBR.

>>>>

Reef depth characteristics were extracted using 10m resolution bathymetry data from the
Great Barrier Reef Marine Park Authority (GBRMPA [-@gbr10gb2020]). For each reef, bathymetry
data (in raster format) were spatially matched with reef polygon features through geometric
intersection analysis. Depth values within each reef polygon were summarized using zonal
statistics to calculate minimum, mean, median, maximum, and standard deviation of depth
values. All depth values were converted to positive values (depths below sea level). The
median depth was used to provide an indicative depth value for each reef.

As bathymetry data are derived from satellite imagery they can be limited in their alignment
with the established GBRMPA reef outlines used in the investigation to characterise reef
locations. Quality control flags were assigned to identify potential data limitations: reefs
with no overlapping bathymetry data, reefs with minimum depth values above sea level
(indicating potential data artifacts), and reefs where bathymetry pixels covered less than
5% of the reef polygon area. These reefs with insufficient data coverage were removed from
analyses (842 reefs of 3806 reefs, or 22%) **\[TI: BG to update\]**.

While data adopted for the given assessment are said to be accurate up to 15m from Mean Sea
Level [@gbr10gb2020], there are many sources of uncertainty in the depth estimates used for
this study. There is potential misalignment between GBRMPA reef outlines, as well as the
usual challenges of deriving data from satellite imagery. In the case of bathymetry,
estimates are subject to different bottom types (rock, sand, seagrass, etc) and how these
affect reflectance, the water penetration depth of satellite sensors, influence of local
conditions (e.g., turbidity and wave activity), and many other factors. The resulting depth
estimates are therefore highly uncertain themselves and are only adopted as a useful
indication rather than an accurate and precise report of reef depths across the GBR.

**\[TI: TODO cite Canonical Reefs repository\]**
